Formal Modelling of Predator Preferences using Molecular Gut-Content Analysis

### 2015-09-17 paper/main.tex
add mark-recapture comment

% !344
Throughout this paper we assume the complete data were generated from underlying Poisson distributions.  Though this assumption is not easily tested, and other factors affecting our model may distort our methods (e.g.\ sampling issues, environmental variables, differential decay rates), we do not believe a better assumption is readily available to handle all the cases we considered.  The Poisson distribution provides reasonable biological interpretation to the available data, data obtained by ecologists interested in prey electivity.  The rate parameters of the Poisson distributions also avoid dependence on estimates of true abundance, a sensitive issue in and of itself.  Alternatively, one could hypothesize a more complex distribution that similarly has good biological interpretation, say a distribution with more than one parameter.  Such a distribution could theoretically perform better than our model when full count data are observed and the sampled data were not generated from a Poisson distribution -- if the data were generated from a Poisson distribution then our likelihood ratio test ensures we are using the most powerful test available.  However, when only unobserved count data is available, any distribution with more than one parameter is unidentifiable.  The Poisson distribution on the other hand allowed us to use a relatively simple EM algorithm to estimate parameters of interest when only unobserved count data are available.  Future development of our methods might consider more complex distributions to handle variables not presently considered in the case of fully observed count data.

Beyond the statistical assumptions, sampling of prey availability should be given careful attention.  To survey prey of interest, we relied on pitfall trapping, which is a reasonable method to quantify potential prey encounters of sit-and-pursue predators, such as wolf spiders \citep{Uetz:1999}.  In general, however, human-made traps are liable to collect prey at rates unequal to that of the predator of interest.  Confounded with researchers' abilities to make sufficiently efficient traps are a host of environmental variables.  Researchers should consider these potential sampling issues and attempt to determine, a priori, how well their traps will characterize prey availability according to the biology of their focal predator.  Mathematically, these issues will directly affect the parameter $c_{st} = \lambda_{st}/\gamma_{st}$.  When human-made traps are relatively more efficient at catching prey than are the predators, $c_{st}$ will be relatively smaller.  Conversely, when the traps are relatively less efficient, $c_{st}$ will be relatively larger.  Unfortunately, these factors can be difficult to control, statistically, practically, or otherwise, and are to some degree unavoidable.

>>>>

% !344
Throughout this paper we assume the complete data were generated from underlying Poisson distributions.  Though this assumption is not easily tested, and other factors affecting our model may distort our methods (e.g.\ sampling issues, environmental variables, differential decay rates), we do not believe a better assumption is readily available to handle all the cases we considered.  The Poisson distribution provides reasonable biological interpretation to the available data, data obtained by ecologists interested in prey electivity.  Alternatively, one could hypothesize a more complex distribution that similarly has good biological interpretation, say a distribution with more than one parameter.  Such a distribution could theoretically perform better than our model when full count data are observed and the sampled data were not generated from a Poisson distribution -- if the data were generated from a Poisson distribution then our likelihood ratio test ensures we are using the most powerful test available.  However, when only unobserved count data is available, any distribution with more than one parameter is unidentifiable.  The Poisson distribution on the other hand allowed us to use a relatively simple EM algorithm to estimate parameters of interest when only unobserved count data are available.  Future development of our methods might consider more complex distributions to handle variables not presently considered in the case of fully observed count data.

Beyond the statistical assumptions, sampling of prey availability should be given careful attention.  To survey prey of interest, we relied on pitfall trapping, which is a reasonable method to quantify potential prey encounters of sit-and-pursue predators, such as wolf spiders \citep{Uetz:1999}.  In general, however, human-made traps are liable to collect prey at rates unequal to that of the predator of interest.  Confounded with researchers' abilities to make sufficiently efficient traps are a host of environmental variables.  Researchers should consider these potential sampling issues and attempt to determine, a priori, how well their traps will characterize prey availability according to the biology of their focal predator.  Mathematically, these issues will directly affect the parameter $c_{st} = \lambda_{st}/\gamma_{st}$.  When human-made traps are relatively more efficient at catching prey than are the predators, $c_{st}$ will be relatively smaller.  Conversely, when the traps are relatively less efficient, $c_{st}$ will be relatively larger.  Unfortunately, these factors can be difficult to control, statistically, practically, or otherwise, and are to some degree unavoidable.  Mark-recapture methods might be implemented to adjust for trap efficiency and estimate the true density of each prey species.  However, recapture rates are likely to be low and this type of data is beyond what is regularly collected in these types of experiments.
